Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-03-11 article/article.Rmd
+ fixes to article

## Methods
### Study system
We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for methodological details]. This data set was particularly well suited for our study, because it provides spatially explicit genomic data for multiple species with vastly different demographic characteristics and evolutionary histories across a broad geographic area with large environmental gradients. As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx. 21' 22.3 km $\times$ 25 km; Figure S1). They visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species. Thus for each second grid cell, for each of the species present in the grid cell, the data set contains information describing the genomic variation of individuals in the cell. We used these grid cells as planning units to develop prioritizations. To reduce computational effort, we only used grid cells that contained at least a single individual ($n = `r nrow(grid.DF)`$). The majority of our analyses were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). To describe the geographic location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates. To describe the environmental characteristics of each grid cell, we obtained 19 bioclimatic variables [approx. 1 km resolution; @r33]. These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). The bioclimatic variables were clipped to the extent of the grid and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first two principal components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers. We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2). To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit (obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472).

@@

First, we grouped individuals into genetic lineages for each species (using a methodology similar to that implemented in \texttt{prabclust R} package; @r479). The purpose of this step was to avoid falsely classifying loci as adaptive. For each species, we ordinated the genomic data and fit models to the ordinations to fit to turn group individuals into populations. Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package] using Gower distances [@r464; using the $\texttt{cluster R}$ package to accommodate sparsity; @r457] to transform the binary loci data into continuous variables (Table S2). To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). We then fit a suite of Gaussian mixture to the ordinations to estimate the probability that individuals belonged to different populations [assuming different covariance structures implemented in the \texttt{mclust} R package; @r477; @r476]. For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G}= 1--`r max(mclust.params.LST[[MODE]]$G)`). We identified the most parsimonious model for each species (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used them to group individuals to populations (Figure S5).

We identified the loci under selection (Table S3). Since outlier locus detection methods compare allele frequencies between populations, we were unable to examine species that contained a single population. For these species ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$), we assumed that all loci were neutral. For species with more than a single population, we ran $\texttt{Bayescan}$ [version 2.1; @r456; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations]. To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the user manual) and assumed they were neutral. Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $\l `r mclust.params.LST[[MODE]]$probability.threshold`$). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

>>>>

## Methods
### Study system
We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for methodological details]. This data set was particularly well suited for our study, because it provides spatially explicit genomic data for multiple species with vastly different demographic characteristics and evolutionary histories across a broad geographic area with large environmental gradients. As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). They visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species. Thus for each second grid cell, for each of the species present in the grid cell, the data set contains information describing the genomic variation of individuals in the cell. We used these grid cells as planning units to develop prioritizations. To reduce computational effort, we only used grid cells that contained at least a single individual ($n = `r nrow(grid.DF)`$). The majority of our analyses were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). To describe the geographic location of each grid cell, we projected the grid into an equi-distant coordinate system ($\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates. To describe the environmental characteristics of each grid cell, we obtained 19 bioclimatic variables [approx. 1 km resolution; @r33]. These layers and the grid cells were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). The bioclimatic variables were clipped to the extent of the grid and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first two principal components (PCs) cumulatively explained `r round(pca.DF[[4]][2],1)` % of the variation and were used to generate two new layers. We calculated the average value for each PC layer in each grid cell, and used these values to characterize the main themes of climatic variation among the grid cells (Figure S2). To examine if the influence of opportunity cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472].

@@

First, we grouped individuals into genetic lineages for each species [using a methodology similar to that implemented in \texttt{prabclust R} package; @r479]. The purpose of this step was to avoid falsely classifying loci as adaptive. For each species, we ordinated the genomic data and fit models to the ordinations to fit to turn group individuals into populations. Specifically, we used non-metric multi-dimensional scaling [NMDS; implemented in the $\texttt{vegan R}$ package; @r458] using Gower distances [@r464; using the $\texttt{cluster R}$ package to accommodate sparsity; @r457] to transform the binary loci data into continuous variables (Table S2). To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). We then fit a suite of Gaussian mixture to the ordinations to estimate the probability that individuals belonged to different populations [assuming different covariance structures implemented in the \texttt{mclust} R package; @r477; @r476]. For each parametrization, we fit a set of models assuming varying numbers of total populations (\textit{G}= 1--`r max(mclust.params.LST[[MODE]]$G)`). We identified the most parsimonious model for each species (lowest Bayesian Information Criterion $BIC$ statistic; Figures S4), and used them to group individuals to populations (Figure S5).

We identified the loci under selection (Table S3). Since outlier locus detection methods compare allele frequencies between populations, we were unable to examine species that contained a single population. For these species ($n = `r sum(sapply(spp.Mclust.LST, function(x) return(ncol(x$mclust$z)==1)))`$), we assumed that all loci were neutral. For species with more than a single population, we ran $\texttt{Bayescan}$ [version 2.1; 1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456]. To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.05$ (following guidelines in the user manual) and assumed they were neutral. Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $\l `r mclust.params.LST[[MODE]]$probability.threshold`$). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate to ensure convergence [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

### 2016-03-11 article/figures.Rmd
+ fixes to article

```{r, echo=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures. Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation. Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets. Letters indicate significant differences ($P < 0.05$).'}
# prepare data
env.correlation.DF$Surrogate.target <- factor(env.correlation.DF$Surrogate.target)
env.correlation.DF2 <- env.correlation.DF %>%
	group_by(Surrogate.target, Method, Type) %>% 
	summarize(genetic.held=max(genetic.held)+0.1) %>%
	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
	left_join(
		data.frame(
			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
			stringsAsFactors=FALSE
		),
		by='Combination'
	) %>% data.frame()
geo.correlation.DF$Surrogate.target <- factor(geo.correlation.DF$Surrogate.target)
geo.correlation.DF2 <- geo.correlation.DF %>%
	group_by(Surrogate.target, Method, Type) %>% 
	summarize(genetic.held=max(genetic.held)+0.1) %>%
	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
	left_join(
		data.frame(
			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
			stringsAsFactors=FALSE
		),
		by='Combination'
	) %>% data.frame()

>>>>

```{r, echo=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures. Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation. Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets. Letters indicate significant differences ($P < 0.05$).'}
# prepare data
env.correlation.DF$Surrogate.target <- factor(env.correlation.DF$Surrogate.target)
env.correlation.DF2 <- env.correlation.DF %>%
	group_by(Surrogate.target, Method, Type) %>% 
	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
	left_join(
		data.frame(
			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
			stringsAsFactors=FALSE
		),
		by='Combination'
	) %>% data.frame()
geo.correlation.DF$Surrogate.target <- factor(geo.correlation.DF$Surrogate.target)
geo.correlation.DF2 <- geo.correlation.DF %>%
	group_by(Surrogate.target, Method, Type) %>% 
	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
	ungroup() %>% mutate(Combination=as.character(paste0(Surrogate.target,'.',Method,'.',Type))) %>% 
	left_join(
		data.frame(
			letter=unname(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(posthoc.correlation.GLHT)$mcletters$Letters))),
			stringsAsFactors=FALSE
		),
		by='Combination'
	) %>% data.frame()
